Population structure and phenotypic variation of *Sclerotinia sclerotiorum* from dry bean in the United States

### 2017-07-11 doc/manuscript/manuscript.Rmd
Address Sydney's comments

<!-- Most of the previous studies on *S. sclerotiorum* had been conducted in the USA, Canada and Australia but not in Europe. 
In those studies, different dry bean cultivars had been rarely used as  hosts to investigate any host-pathogen relationships or specificity.
No white mold resistant dry bean cultivars exist.
Canopy architecture and growth pattern of some cultivars enable improved disease avoidance [@kolkman2002agronomic]. -->
To date, no commercial white mold-resistant dry bean cultivars exist. 
Field trials in some regions of the US have shown the G122 cultivar is partially resistant while the Beryl cultivar is susceptible [@otto-hanson2011variation].

@@

Sclerotia of *S. sclerotiorum* were collected over several years from grower fields and/or dry bean nurseries in 11 states of the United States of America, Mexico, France, and Australia (Table \@ref(tab:isolate-table)).
The dry bean nurseries in different regions are indicated by wmn (white mold nursery) or mtc (Mitchell) and all other isolates came from commercial fields (Table \@ref(tab:isolate-table)).
> SEE: I was not sure what "dry bean nurseries" were because I'm use to seeing them referred to as the white mold nurseries.  I think if the introduction uses one term for it, this would be acceptable.  That or consider providing a description of the white mold nurseries in the methods section.  This now reminds me.. there is a publication by Lindsey Otto-Hanson and Steadman that gave a good description of these.  In fact, she called them multisite screening locations.  Here's the link: http://apsjournals.apsnet.org/doi/abs/10.1094/PDIS-11-10-0865

The dry bean nurseries were 5m x 10m in size and maintained without applying fungicides to observe natural incidence of white mold.
The nursery plots were incorporated with a basal dressing of N:P:K = 1:3:2 and side dressing of 0:3:2 during the growing season.
Although the nursery locations were the same over sampling years, sampling plots within a location varied for sampling years.
Sampling of diseased tissue was carried out on a 'W' or 'V' shape transect, or randomly.
Because sampling depended on disease incidence, the number of samples isolated varied from year to year.
In total, we collected 366 isolates that were used in the present study.

@@

Each isolate was regenerated by activating one or several sclerotia according to the method described previously [@matheron1989in-vitro].
Once actively growing, mycelial compatibility group (MCG) and aggressiveness was determined for each isolate.
MCG was determined through co-culturing isolates in pairs on potato-dextrose agar (PDA) plates, where incompatibility of different MCG resulted in formation of a mycelial barrier or barrage line [@kohn1990mycelial].
Since MCG are considered mutually exclusive, pairing was not required between all isolates in the study.
> SEE: Is that correct?  Perhaps consult with Becky or Rachana and then elaborate upon the sentence above.

Aggressiveness of each isolate was assessed using a straw test that rated necrotic lesion size on a scale from 1--9 with increasing aggressiveness [@petzoldt1996straw; @teran2006modified].
> SEE: Rating is not clear from the single sentence above.  On what plant host was it evaluated?  In the greenhouse?  What is a rating of 1 compared to a rating of 9 on the straw test?

DNA was purified from 100 mg mycelium collected from an actively growing culture of each isolate using a Qiagen DNeasy Plant Mini Kit (Qiagen Inc., Valencia, CA) according to the manufacturer's instructions.
> SEE: More details on how tissue was disrupted may be needed here.

@@

We genotyped *S. sclerotiorum* isolates using 16 microsatellite primer pairs developed previously [@sirjusingh2001characterisation].
Polymerase chain reaction (PCR) was carried out as described using primers labeled with FAM fluorophore, and resulting PCR products were first resolved in a 1.5% agarose gel stained with ethidium bromide to ensure product was within the expected size range.
> SEE: Pretty rare to see PCR spelled out in full these days. I suggest just using the abbreviation and if PeerJ asks us to spell it out, we will.

Capillary electrophoresis (fragment analysis) of amplicons, with size standard GeneScan™ 500 LIZ®, was performed using an ABI 3730 genetic analyzer (Life Technologies Corporation, Carlsbad, CA) at the Michigan State University Genomic Sequencing Center.

> ZNK: how were the data received and formatted?

> SEE: data were received as .fsa files, which are fasta-type files.  I have the raw data if you want them.  There is, apparently, an R-package called Fragman used for reading the files and can be used for scoring.  Anyway, in this case, I believe Rachana used PeakScanner (free software) to read the files and she recorded the fragment sizes manually (ie. read the chromatogram and wrote down in Excel what size the peak was).

@@

Of the 16 microsatellite loci, five included compound repeats, which made it difficult to accurately/confidently bin alleles into fragment sizes expected for each locus based on the described repeat motif. 
> SEE: I would not say allele scoring was difficult, because scoring an allele that is different than the size expected is really the same difficulty level as one that is at the expected size.  It's the accurate binning of alleles into the expected fragment size that is difficult.

Loci with compound repeats were removed for the reported statistics, but were processed and included in data analyses to ensure the integrity of the results with 11 loci as compared to 16.
We assessed the power of our 11 markers by generating a genotype accumulation curve in the R package *poppr* version 2.4.1, looking for evidence of saturation, which would indicate that loci were sufficiently sampled to adequately represent the full set of genotypes [@arnaud2007standardizing; @kamvar2015novel].
To control for non-independent samples (i.e. samples from the same location, region, year, and plant), data were clone-corrected for subsequent analyses.
> SEE: it is not clear what non-independent samples are.  Although your description is accurate and concise, a more general description might enable the reader to better understand what is meant here.

@@

We used analysis of variance (ANOVA) to assess if MCGs differed in aggressiveness.
To minimize complications due to small sample sizes, we chose the top 10 MCGs, representing 57% of the isolates collected. Between-group differences were assessed using Tukey's HSD test using the `HSD.test()` function in the package *agricolae* version 1.2.4 [@mendiburu2015agricolae].
> SEE: It's not clear what groups were used to compare aggressiveness in the ANOVA.  What alpha-level was used for the cut-off?

@@

## Genetic Diversity {-}
> SEE: deleting this header?

@@

We used analysis of molecular variance (AMOVA) with Bruvo's genetic distance in *poppr* version 2.4.1 to assess differentiation associated with wmn by quantifying the differentiation conferred by source and region [@excoffier1992analysis; @bruvo2004simple; @kamvar2014poppr].
Source in this case was defined as a population belonging to a wmn or not and region defined as state. 
> SEE: is mtc included with "wmn"? At the beginning of the methods, I thought both were considered dry bean nurseries.

To identify specific regions that were more differentiated than others, we used discriminant analysis of principal components (DAPC) as implemented in *adegenet* version 2.1.0, assessing the per-sample posterior group assignment probability [@jombart2008adegenet]. 
This method decomposed the genetic data into principal components, and then used a subset of these as the inputs for discriminant analysis, which attempts to minimize within-group variation and maximize among-group variation [@jombart2010discriminant]. 
To avoid over-fitting the data, the optimal number of principal components was selected by using the *adegenet* function `xvalDapc()`. 
This function implemented a cross-validation procedure to iterate over an increasing number of principal components on a subset (90%) of the data, trying to find the minimum number of principal components that maximized the rate of successful group reassignment. 
We additionally assessed genotypic diversity by calculating Stoddart and Taylor's index (*G*) [@stoddart1988genotypic], Shannon's index (*H*) [@shannon2001mathematical], Simpson's index ($\lambda$) [@simpson1949measurement], and evenness ($E_5$, *eMLG*) [@heck1975explicit; @hurlbert1971nonconcept; @pielou1975ecological; @grunwald2003analysis]. 
To assess the potential for random mating, we tested for linkage disequilibrium with the index of association, $I_A$ and its standardized version, $\bar{r}_d$ using 999 permutations [@agapow2001indices; @smith1993how; @brown1980multilocus].
Both genotypic diversity and linkage disequilibrium were calculated in *poppr* version 2.4.1 [@kamvar2014poppr].

@@

As many disparate geographic regions shared multilocus genotypes, we wanted to know if we could find any patterns in connectivity. 
> SEE: avoid beginning a paragraph with a preposition because it forces the reader to read and remember the first half of the sentence before understanding it by reading the second half of the sentence.  It's better to reorganize and start with the subject, such as "Connectivity between shared multilocus genotypes identified in disparate geographic regions was evaluated."

To assess the connectivity of shared genotypes across regions, we first tabulated the multilocus genotypes shared between at least two populations (defined as states or countries) with the *poppr* function `mlg.crosspop()` [@kamvar2014poppr].
From these data, we constructed a graph with populations as nodes and shared genotypes as edges (connections) between nodes using the R packages *igraph* version 1.0.1 [@csardi2006igraph], *dplyr* version 0.7.1 [@dplyr], and *purrr* version 0.2.2.2 [@purrr].
Each node was weighted by the fraction of shared MLGs.
Each edge represented a single MLG, but because a single MLG could cross populations, that MLG could be represented by (n\*(n-1))/2 edges where n represented the number of populations it crossed.
Edges were weighted by $1-P_{sex}$, where $P_{sex}$ was the probability of encountering the same genotype via two independent meiotic events [@arnaud2007standardizing; @parks1993study].
We then used the `cluster_optimal()` function from *igraph* to define communities [@csardi2006igraph].
The graph was plotted using the R packages *ggplot2* version 2.2.1 [@ggplot2] and *ggraph* version 1.0.0 [@ggraph].
To ensure that we captured the same community signal, we additionally performed this analysis including the 5 polymorphic markers described above.

@@

Contrary to above results, @aldrich-wolfe2015genetic who did a survey on *S. sclerotiorum* from different host crops in the North Central states of the United States found a single microsatellite haplotype to be prevalent throughout the region.
A study in the UK that examined this pathogen from 12 populations in different crop plants and weed species *Ranunculus acris* (meadow buttercup) also reported one microsatellite haplotype was widely distributed across hosts and areas [@clarkson2013population].
Reason for our isolates from dry bean not to show a widespread MLG may be due to extensive sexual outcrossing and movement sclerotia by human activities.
Though conditions of outcrossing for *S. sclerotiorum* is not well studied, it is believed weather and agricultural conditions play a role in production and release of ascospores of this pathogen [@huang1991induction; @sexton2006population].
One study found release of spores from apothecia declined over 36- 168 H at low RH of 65-75% but not at high RH of 90-95% [@clarkson2003ascospore].
The same study reported increased ascospores discharge at 20-25$^{\circ}$ C compared to 15$^{\circ}$ C and the survival of *S. sclerotiorum* ascospores was reduced at high temperature and RH.
Our samples were collected from 2003 to 2012 and except for the dry bean nurseries, came from different fields.
Different agro-ecological conditions prevailed during sampling seasons may have resulted in different rates of sexual outcrossing of the pathogen and hence reduced the chance of extracting a common genotype.
When conditions are favorable for ascospore production, release, and survival, co-infection of a single site on a plant by multiple ascospores is possible [@sexton2006population].
This may eventually result in facilitating outcrossing and recombination.

>>>>

<!-- Most of the previous studies on *S. sclerotiorum* had been conducted in the USA, Canada and Australia but not in Europe. 
In those studies, different dry bean cultivars had been rarely used as  hosts to investigate any host-pathogen relationships or specificity.
No white mold resistant dry bean cultivars exist.
Canopy architecture and growth pattern of some cultivars enable improved disease avoidance [@kolkman2002agronomic]. -->
To date, no commercial white mold-resistant dry bean cultivars exist. 
Field trials in some regions of the US have shown the G122 cultivar is partially resistant while the Beryl cultivar is susceptible [@otto-hanson2011variation; @steadman2004identification].

@@

Sclerotia of *S. sclerotiorum* were collected over several years from grower fields and/or white mold nurseries in 11 states of the United States of America, Mexico, France, and Australia (Table \@ref(tab:isolate-table)).
The white mold nurseries in different regions are indicated by wmn (white mold nursery) or mtc (Mitchell) and all other isolates came from commercial fields (Table \@ref(tab:isolate-table)).

> SEE: I was not sure what "dry bean nurseries" were because I'm use to seeing them referred to as the white mold nurseries.  I think if the introduction uses one term for it, this would be acceptable.  That or consider providing a description of the white mold nurseries in the methods section.  This now reminds me.. there is a publication by Lindsey Otto-Hanson and Steadman that gave a good description of these.  In fact, she called them multisite screening locations.  Here's the link: http://apsjournals.apsnet.org/doi/abs/10.1094/PDIS-11-10-0865

The white mold nurseries were 5m x 10m in size and maintained without applying fungicides to observe natural incidence of white mold.
The nursery plots were incorporated with a basal dressing of N:P:K = 1:3:2 and side dressing of 0:3:2 during the growing season.
Although the nursery locations were the same over sampling years, sampling plots within a location varied for sampling years.
Sampling of diseased tissue was carried out on a 'W' or 'V' shape transect, or randomly.
Because sampling depended on disease incidence, the number of samples isolated varied from year to year.
In total, we collected 366 isolates that were used in the present study.

@@

Each isolate was regenerated by activating one or several sclerotia according to the method described previously [@matheron1989in-vitro].
Once actively growing, mycelial compatibility group (MCG) and aggressiveness was determined for each isolate.
MCG was determined through co-culturing isolates in pairs on potato-dextrose agar (PDA) plates, where incompatibility of different MCG resulted in formation of a mycelial barrier or barrage line [@kohn1990mycelial].
Since MCG are considered mutually exclusive, pairing was not required between all isolates in the study.

> SEE: Is that correct?  Perhaps consult with Becky or Rachana and then elaborate upon the sentence above.

Aggressiveness of each isolate was assessed using a straw test that rated necrotic lesion size on a scale from 1--9 with increasing aggressiveness [@petzoldt1996straw; @teran2006modified].

> SEE: Rating is not clear from the single sentence above.  On what plant host was it evaluated?  In the greenhouse?  What is a rating of 1 compared to a rating of 9 on the straw test?

DNA was purified from 100 mg mycelium collected from an actively growing culture of each isolate using a Qiagen DNeasy Plant Mini Kit (Qiagen Inc., Valencia, CA) according to the manufacturer's instructions.

> SEE: More details on how tissue was disrupted may be needed here.

@@

We genotyped *S. sclerotiorum* isolates using 16 microsatellite primer pairs developed previously [@sirjusingh2001characterisation].
PCR was carried out as described using primers labeled with FAM fluorophore, and resulting PCR products were first resolved in a 1.5% agarose gel stained with ethidium bromide to ensure product was within the expected size range.

Capillary electrophoresis (fragment analysis) of amplicons, with size standard GeneScan™ 500 LIZ®, was performed using an ABI 3730 genetic analyzer (Life Technologies Corporation, Carlsbad, CA) at the Michigan State University Genomic Sequencing Center. 
Alleles were binned using PeakScanner version 1.0 (Life Technologies Corporation, Carlsbad, CA) and recorded manually in an excel sheet.

@@

Of the 16 microsatellite loci, five included compound repeats, which made it challenging to accurately/confidently bin alleles into fragment sizes expected for each locus based on the described repeat motif. 
Loci with compound repeats were removed for the reported statistics, but were processed and included in data analyses to ensure the integrity of the results with 11 loci as compared to 16.
We assessed the power of our 11 markers by generating a genotype accumulation curve in the R package *poppr* version 2.4.1, looking for evidence of saturation, which would indicate that loci were sufficiently sampled to adequately represent the full set of genotypes [@arnaud2007standardizing; @kamvar2015novel].
To avoid including samples that were potentially collected from the same plant (which increases the probability of collecting the same sample more than once), data were clone-corrected on the region/source/host/year hierarchy for subsequent analyses.

@@

We used analysis of variance (ANOVA) to assess if MCGs differed in aggressiveness.
To minimize complications due to small sample sizes, we chose the top 10 MCGs, representing 57% of the isolates collected. Between-MCG differences were assessed using Tukey's HSD test with an alpha of 0.05 using the `HSD.test()` function in the package *agricolae* version 1.2.4 [@mendiburu2015agricolae].

@@

## Genetic Diversity {-}

@@

We used analysis of molecular variance (AMOVA) with Bruvo's genetic distance in *poppr* version 2.4.1 to assess differentiation associated with wmn by quantifying the differentiation conferred by source and region [@excoffier1992analysis; @bruvo2004simple; @kamvar2014poppr].
Source in this case was defined as a population belonging to a wmn or not and region defined as state.

To identify specific regions that were more differentiated than others, we used discriminant analysis of principal components (DAPC) as implemented in *adegenet* version 2.1.0, assessing the per-sample posterior group assignment probability [@jombart2008adegenet]. 
This method decomposes the genetic data into principal components, and then uses a subset of these as the inputs for discriminant analysis, which attempts to minimize within-group variation and maximize among-group variation [@jombart2010discriminant]. 
To avoid over-fitting the data, the optimal number of principal components was selected by using the *adegenet* function `xvalDapc()`. 
This function implemented a cross-validation procedure to iterate over an increasing number of principal components on a subset (90%) of the data, trying to find the minimum number of principal components that maximized the rate of successful group reassignment. 
We additionally assessed genotypic diversity by calculating Stoddart and Taylor's index (*G*) [@stoddart1988genotypic], Shannon's index (*H*) [@shannon2001mathematical], Simpson's index ($\lambda$) [@simpson1949measurement], and evenness ($E_5$, *eMLG*) [@heck1975explicit; @hurlbert1971nonconcept; @pielou1975ecological; @grunwald2003analysis]. 
To assess the potential for random mating, we tested for linkage disequilibrium with the index of association, $I_A$ and its standardized version, $\bar{r}_d$ using 999 permutations [@agapow2001indices; @smith1993how; @brown1980multilocus].
Both genotypic diversity and linkage disequilibrium were calculated in *poppr* version 2.4.1 [@kamvar2014poppr].

@@

We wanted to evaluate patterns of connectivity between shared multilocus genotypes across geographic regions.
We first tabulated the multilocus genotypes shared between at least two populations (defined as states or countries) with the *poppr* function `mlg.crosspop()` [@kamvar2014poppr].
From these data, we constructed a graph with populations as nodes and shared genotypes as edges (connections) between nodes using the R packages *igraph* version 1.0.1 [@csardi2006igraph], *dplyr* version 0.7.1 [@dplyr], and *purrr* version 0.2.2.2 [@purrr].
Each node was weighted by the fraction of shared MLGs.
Each edge represented a single MLG, but because a single MLG could cross populations, that MLG could be represented by (n\*(n-1))/2 edges where n represented the number of populations it crossed.
Edges were weighted by $1-P_{sex}$, where $P_{sex}$ is the probability of encountering the same genotype via two independent meiotic events [@arnaud2007standardizing; @parks1993study].
We then used the `cluster_optimal()` function from *igraph* to define communities [@csardi2006igraph].
The graph was plotted using the R packages *ggplot2* version 2.2.1 [@ggplot2] and *ggraph* version 1.0.0 [@ggraph].
To ensure that we captured the same community signal, we additionally performed this analysis including the 5 polymorphic markers described above.

@@

Contrary to above results, @aldrich-wolfe2015genetic who did a survey on *S. sclerotiorum* from different host crops in the North Central states of the United States found a single microsatellite haplotype to be prevalent throughout the region.
A study in the UK that examined this pathogen from 12 populations in different crop plants and weed species *Ranunculus acris* (meadow buttercup) also reported one microsatellite haplotype was widely distributed across hosts and areas [@clarkson2013population].
Reason for our isolates from dry bean not to show a widespread MLG may be due to extensive sexual outcrossing and movement sclerotia by human activities.
Though conditions of outcrossing for *S. sclerotiorum* is not well studied, it is believed weather and agricultural conditions play a role in production and release of ascospores of this pathogen [@huang1991induction; @sexton2006population].
One study found release of spores from apothecia declined over 36- 168 H at low RH of 65-75% but not at high RH of 90-95% [@clarkson2003ascospore].
The same study reported increased ascospores discharge at 20-25$^{\circ}$ C compared to 15$^{\circ}$ C and the survival of *S. sclerotiorum* ascospores was reduced at high temperature and RH.
Our samples were collected from 2003 to 2012 and except for the white mold nurseries, came from different fields.
Different agro-ecological conditions prevailed during sampling seasons may have resulted in different rates of sexual outcrossing of the pathogen and hence reduced the chance of extracting a common genotype.
When conditions are favorable for ascospore production, release, and survival, co-infection of a single site on a plant by multiple ascospores is possible [@sexton2006population].
This may eventually result in facilitating outcrossing and recombination.
